---
title: "Linear_mixed_models_functional_groups"
author: "Hannah Sørine Gerlich"
format: html
editor: visual
message: false
warning: false
output: false
---

# Analyses of phenological responses to snowmelt timing and temperature among functional groups

This file contains all the code necessary to perform linear mixed modelling on the effect of temperature and snowmelt on phenology among functional groups and arthropods with specific traits.

Initially, we load all the relevant packages.

```{r}
pkgs <- c("tidyverse", "readxl", "stats", "lme4", "lmerTest", "car",
          "sjPlot", "MASS", "gtsummary")

vapply(pkgs, library, FUN.VALUE = logical(1L), character.only = TRUE,
       logical.return = TRUE)
```

```{r}
df_phen_event <-
  read.csv(
    "Data/Pheno_air_temp.csv",
    sep = ",",
    stringsAsFactors = FALSE,
    header = TRUE
  )

dfsnowmelt_climatestation <-
  readxl::read_xlsx("Data/Snowmelt_climatestation.xlsx")

df_phen_event$Snowmelt <-
  dfsnowmelt_climatestation$DOY[match(paste0(df_phen_event$Year),
                                      paste0(dfsnowmelt_climatestation$Year))]
```

Assign functional groups and traits.

```{r}
df_phen_event %>%
  mutate(Order = case_when(
    SpeciesID == "Acari" ~ "Decomposer",
    SpeciesID == "ANMU" ~ "Mixed feeder",
    SpeciesID == "Aphidoidea" ~ "Herbivor",
    SpeciesID == "Chalcidoidea" ~ "Parasitoid",
    SpeciesID == "CHCE" ~ "Mixed feeder",
    SpeciesID == "Coccoidea" ~ "Herbivor",
    SpeciesID == "Collembola" ~ "Decomposer",
    SpeciesID == "Culicidae" ~ "Mixed feeder",
    SpeciesID == "Ichneumonidae" ~ "Parasitoid",
    SpeciesID == "Linyphiidae" ~ "Predator",
    SpeciesID == "Lycosidae" ~ "Predator",
    SpeciesID == "MYSC" ~ "Mixed feeder",
    SpeciesID == "Nymphalidae" ~ "Mixed feeder",
    SpeciesID == "Phoridae" ~ "Mixed feeder",
    SpeciesID == "Scathophagidae" ~ "Mixed feeder",
    SpeciesID == "Thomisidae" ~ "Predator")) -> df_phen_event

df_phen_event %>%
  mutate(Emerge = case_when(
    SpeciesID == "Acari" ~ "Early",
    SpeciesID == "ANMU" ~ "Late",
    SpeciesID == "Aphidoidea" ~ "Late",
    SpeciesID == "Chalcidoidea" ~ "Late",
    SpeciesID == "CHCE" ~ "Early",
    SpeciesID == "Coccoidea" ~ "Late",
    SpeciesID == "Collembola" ~ "Early",
    SpeciesID == "Culicidae" ~ "Late",
    SpeciesID == "Ichneumonidae" ~ "Late",
    SpeciesID == "Linyphiidae" ~ "Early",
    SpeciesID == "Lycosidae" ~ "Early",
    SpeciesID == "MYSC" ~ "Early",
    SpeciesID == "Nymphalidae" ~ "Late",
    SpeciesID == "Phoridae" ~ "Late",
    SpeciesID == "Scathophagidae" ~ "Late",
    SpeciesID == "Thomisidae" ~ "Early")) -> df_phen_event

df_phen_event %>%
  mutate(Trait = case_when(
    SpeciesID == "Acari" ~ "Surface",
    SpeciesID == "ANMU" ~ "Flying",
    SpeciesID == "Aphidoidea" ~ "Surface",
    SpeciesID == "Chalcidoidea" ~ "Flying",
    SpeciesID == "CHCE" ~ "Flying",
    SpeciesID == "Coccoidea" ~ "Surface",
    SpeciesID == "Collembola" ~ "Surface",
    SpeciesID == "Culicidae" ~ "Flying",
    SpeciesID == "Ichneumonidae" ~ "Flying",
    SpeciesID == "Linyphiidae" ~ "Surface",
    SpeciesID == "Lycosidae" ~ "Surface",
    SpeciesID == "MYSC" ~ "Flying",
    SpeciesID == "Nymphalidae" ~ "Flying",
    SpeciesID == "Phoridae" ~ "Flying",
    SpeciesID == "Scathophagidae" ~ "Flying",
    SpeciesID == "Thomisidae" ~ "Surface")) -> df_phen_event
```

We find strong collinearity between temperature and snowmelt for families in the snowbed plot (Art6) and Nymphalidae in one mesic heath plot (Art3). We also do not consider phenology observations for spiders (predators) and decomposers in the window trap (Plot 1) reliable. Therefore, we have to remove these observations from the data.

```{r}
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art1" &
                    (df_phen_event$Order == "Decomposer")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art1" &
                    (df_phen_event$Order == "Predator")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art3" &
                    (df_phen_event$SpeciesID == "Nymphalidae")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$SpeciesID == "Lygaeidae"), ]

df_phen_event <- subset(df_phen_event,!Plot == "Art6")
```

Assign habitat types.

```{r}
df_phen_event %>%
  mutate(
    Habitat = case_when(
      Plot == "Art1" ~ "Pond",
      Plot == "Art2" ~ "Wet fen",
      Plot == "Art3" ~ "Mesic heath",
      Plot == "Art4" ~ "Mesic heath",
      Plot == "Art5" ~ "Arid heath",
      Plot == "Art7" ~ "Arid heath"
    )
  ) -> df_phen_event
```

Remove taxa in the dataset for which we do not have enough data. These omly contain NA values.

```{r}
df_phen_event$Order[is.na(df_phen_event$Order)] <- "NA"
df_phen_event <- subset(df_phen_event, Order != "NA")
```

Because the temperature variable is specific for each taxon-by-plot combination, we need to account for within subject variation. Other than that, we also have to add year as a covariate to detrend the model.

```{r}
df_phen_event$speciesplot <-
  as.factor(paste(df_phen_event$SpeciesID, df_phen_event$Plot))

df_phen_event$Peak_Temp_new <-
  df_phen_event$Peak_Temp30 - tapply(df_phen_event$Peak_Temp30, df_phen_event$speciesplot, mean)[df_phen_event$speciesplot]
df_phen_event$Onset_Temp_new <-
  df_phen_event$Onset_Temp30 - tapply(df_phen_event$Onset_Temp30, df_phen_event$speciesplot, mean)[df_phen_event$speciesplot]
df_phen_event$End_Temp_new <-
  df_phen_event$End_Temp30 - tapply(df_phen_event$End_Temp30, df_phen_event$speciesplot, mean)[df_phen_event$speciesplot]
```

First, we test differences in **peak** phenological responses to climate variables among functional groups.

## Peak phenology

First, decomposer is compared to the other functional groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
#contrasts(df_phen_event$Plot) <- contr.treatment(7, base = 1)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 1)


model_decomp <- lmer(Peak ~ 1 + Snowmelt*Order + Peak_Temp_new*Order + Year + (1 + Peak_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data=df_phen_event) 

summary(model_decomp)
anova(model_decomp)
```

```{r output=TRUE}
tab_model(model_decomp, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Peak_Temp_new", "Year", "Order2", "Order3", "Order4", "Order5", "Snowmelt:Order2",  "Snowmelt:Order3",  "Snowmelt:Order4", "Snowmelt:Order5", "Order2:Peak_Temp_new",  "Order3:Peak_Temp_new",  "Order4:Peak_Temp_new", "Order5:Peak_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Herbivore", "Parasitoid", "Mixed feeder", "Predator",
                         "Temp", "Year", "Snowmelt:Herbivore", "Snowmelt:Parasitoid", "Snowmelt:Mixed feeder", "Snowmelt:Predator", "Temp:Herbivore", "Temp:Parasitoid", "Temp:Mixed feeder", "Temp:Predator"), title = "Peak ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Decomposer", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Herbivore is compared to the other functional groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 2)


model_herb <- lmer(Peak ~ 1 + Snowmelt*Order + Peak_Temp_new*Order + Year + (1 + Peak_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data=df_phen_event) 
summary(model_herb)
```

```{r output=TRUE}
tab_model(model_herb, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Peak_Temp_new", "Year", "Order1", "Order3", "Order4", "Order5", "Snowmelt:Order1",  "Snowmelt:Order3",  "Snowmelt:Order4", "Snowmelt:Order5", "Order1:Peak_Temp_new",  "Order3:Peak_Temp_new",  "Order4:Peak_Temp_new", "Order5:Peak_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Parasitoid", "Mixed feeder", "Predator",
                         "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Parasitoid", "Snowmelt:Mixed feeder", "Snowmelt:Predator", "Temp:Decomposer", "Temp:Parasitoid", "Temp:Mixed feeder", "Temp:Predator"), title = "Peak ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Herbivore", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Parasitoid is compared to the other functional groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 3)


model_para <- lmer(Peak ~ 1 + Snowmelt*Order + Peak_Temp_new*Order + Year + (1 + Peak_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data=df_phen_event)
summary(model_para)
```

```{r output=TRUE}
tab_model(model_para, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)",  "Snowmelt", "Peak_Temp_new", "Year", "Order1", "Order2", "Order4", "Order5", "Snowmelt:Order1",  "Snowmelt:Order2",  "Snowmelt:Order4", "Snowmelt:Order5", "Order1:Peak_Temp_new",  "Order2:Peak_Temp_new",  "Order4:Peak_Temp_new", "Order5:Peak_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Herbivore", "Mixed feeder", "Predator",
                         "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Herbivore", "Snowmelt:Mixed feeder", "Snowmelt:Predator", "Temp:Decomposer", "Temp:Herbivore", "Temp:Mixed feeder", "Temp:Predator"), title = "Peak ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Parasitoids", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Mixed feeders compared with other functional groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 4)

model_poll <- lmer(Peak ~ 1 + Snowmelt*Order + Peak_Temp_new*Order + Year + (1 + Peak_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data=df_phen_event)
summary(model_poll)
```

```{r output=TRUE}
tab_model(model_poll, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE, show.re.var = TRUE,
          terms = c("(Intercept)",  "Snowmelt", "Peak_Temp_new", "Year", "Order1", "Order2", "Order3", "Order5", "Snowmelt:Order1",  "Snowmelt:Order2",  "Snowmelt:Order3", "Snowmelt:Order5", "Order1:Peak_Temp_new",  "Order2:Peak_Temp_new",  "Order3:Peak_Temp_new", "Order5:Peak_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Herbivore", "Parasitoid", "Predator",
                         "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Herbivore", "Snowmelt:Parasitoid", "Snowmelt:Predator", "Temp:Decomposer", "Temp:Herbivore", "Temp:Parasitoid", "Temp:Predator"), title = "Peak ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Mixed feeder", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Predators compared to other functional groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 5)


model_pred <- lmer(Peak ~ 1 + Snowmelt*Order + Peak_Temp_new*Order + Year + (1 + Peak_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data=df_phen_event)
```

```{r output=TRUE}
tab_model(model_pred, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)",  "Snowmelt", "Peak_Temp_new", "Year", "Order1", "Order2", "Order3", "Order4", "Snowmelt:Order1",  "Snowmelt:Order2",  "Snowmelt:Order3", "Snowmelt:Order4", "Order1:Peak_Temp_new",  "Order2:Peak_Temp_new",  "Order3:Peak_Temp_new", "Order4:Peak_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Herbivore", "Parasitoid", "Mixed feeder",
                         "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Herbivore", "Snowmelt:Parasitoid", "Snowmelt:Mixed feeder", "Temp:Decomposer", "Temp:Herbivore", "Temp:Parasitoid", "Temp:Mixed feeder"), title = "Peak ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Predators", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

## Duration of activity

Decomposer compared to other groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 1)

model_decomp <- lmer(Duration ~ 1 + Snowmelt*Order + Onset_Temp_new*Order + Year + (1 + Onset_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data=df_phen_event)
summary(model_decomp)
```

```{r output=TRUE}
tab_model(model_decomp, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Onset_Temp_new", "Year", "Order2", "Order3", "Order4", "Order5", "Snowmelt:Order2",  "Snowmelt:Order3",  "Snowmelt:Order4", "Snowmelt:Order5", "Order2:Onset_Temp_new",  "Order3:Onset_Temp_new",  "Order4:Onset_Temp_new", "Order5:Onset_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Herbivore", "Parasitoid", "Mixed feeder", "Predator",
                         "Temp", "Year", "Snowmelt:Herbivore", "Snowmelt:Parasitoid", "Snowmelt:Mixed feeder", "Snowmelt:Predator", "Temp:Herbivore", "Temp:Parasitoid", "Temp:Mixed feeder", "Temp:Predator"), title = "Duration ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Decomposer", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Herbivore compared to other groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 2)

model_herb <- lmer(Duration ~ 1 + Snowmelt*Order + Onset_Temp_new*Order + Year + (1 + Onset_Temp_new | speciesplot) + (1 + Snowmelt | speciesplot), data = df_phen_event)
summary(model_herb)
```

```{r output=TRUE}
tab_model(model_herb, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Onset_Temp_new", "Year", "Order1", "Order3", "Order4", "Order5", "Snowmelt:Order1",  "Snowmelt:Order3",  "Snowmelt:Order4", "Snowmelt:Order5", "Order1:Onset_Temp_new",  "Order3:Onset_Temp_new",  "Order4:Onset_Temp_new", "Order5:Onset_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Parasitoid", "Mixed feeder", "Predator",
                         "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Parasitoid", "Snowmelt:Mixed feeder", "Snowmelt:Predator", "Temp:Decomposer", "Temp:Parasitoid", "Temp:Mixed feeder", "Temp:Predator"), title = "Duration ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Herbivores", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Parasitoids compared to other groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 3)


model_para <- lmer(Duration ~ 1 + Snowmelt*Order + Onset_Temp_new*Order + Year + (1 | Year) + (1 | SpeciesID:Plot) + (1 + Snowmelt + Onset_Temp_new | SpeciesID:Plot), data=df_phen_event)
summary(model_para)
```

```{r output=TRUE}
tab_model(model_para, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Year", "Onset_Temp_new", "Order1", "Order2", "Order4", "Order5", "Snowmelt:Order1",  "Snowmelt:Order2",  "Snowmelt:Order4", "Snowmelt:Order5", "Order1:Onset_Temp_new",  "Order2:Onset_Temp_new",  "Order4:Onset_Temp_new", "Order5:Onset_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Herbivore", "Mixed feeder", "Predator", "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Herbivore", "Snowmelt:Mixed feeder", "Snowmelt:Predator", "Temp:Decomposer", "Temp:Herbivore", "Temp:Mixed feeder", "Temp:Predator"), title = "Duration ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Parasitoids", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Mixed feeders compared to other groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 4)

model_poll <- lmer(Duration ~ 1 + Snowmelt*Order + Onset_Temp_new*Order + Year + (1 | Year) + (1 | SpeciesID:Plot) + (1 + Snowmelt + Onset_Temp_new | SpeciesID:Plot), data=df_phen_event)

summary(model_poll)
```

```{r output=TRUE}
tab_model(model_poll, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Year", "Onset_Temp_new", "Order1", "Order2", "Order3", "Order5", "Snowmelt:Order1",  "Snowmelt:Order2",  "Snowmelt:Order3", "Snowmelt:Order5", "Order1:Onset_Temp_new",  "Order2:Onset_Temp_new",  "Order3:Onset_Temp_new", "Order5:Onset_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Herbivore", "Parasitoid", "Predator", "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Herbivore", "Snowmelt:Parasitoid", "Snowmelt:Predator", "Temp:Decomposer", "Temp:Herbivore", "Temp:Parasitoid", "Temp:Predator"), title = "Duration ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Mixed feeder", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Predators compared to other groups.

```{r}
df_phen_event$Order <- as.factor(df_phen_event$Order)
contrasts(df_phen_event$Order) <- contr.treatment(5, base = 5)

model_pred <- lmer(Duration ~ 1 + Snowmelt*Order + Onset_Temp_new*Order + Year + (1 | Year) + (1 | SpeciesID:Plot) + (1 + Snowmelt + Onset_Temp_new | SpeciesID:Plot), data=df_phen_event)
summary(model_pred)
```

```{r output=TRUE}
tab_model(model_pred, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
          terms = c("(Intercept)", "Snowmelt", "Year", "Onset_Temp_new", "Order1", "Order2", "Order3", "Order4", "Snowmelt:Order1",  "Snowmelt:Order2",  "Snowmelt:Order3", "Snowmelt:Order4", "Order1:Onset_Temp_new",  "Order2:Onset_Temp_new",  "Order3:Onset_Temp_new", "Order4:Onset_Temp_new"),
         pred.labels = c("Intercept", "Snowmelt", "Decomposer", "Herbivore", "Parasitoid", "Mixed feeder", "Temp", "Year", "Snowmelt:Decomposer", "Snowmelt:Herbivore", "Snowmelt:Parasitoid", "Snowmelt:Mixed feeder", "Temp:Decomposer", "Temp:Herbivore", "Temp:Parasitoid", "Temp:Mixed feeder"), title = "Duration ~ Snowmelt*Group + Temp*Group + Year + (1 + Temp + Snowmelt | Taxa:Plot)", dv.labels = "Compared group: Predators", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

## Phenological responses of arthropod taxa with different traits

Here we test if arthropods with different traits (early/late active, flying/surface dwelling) respond in the same way.

```{r}
df_phen_event$Emerge <- as.factor(df_phen_event$Emerge)
df_phen_event$Trait <- as.factor(df_phen_event$Trait)
contrasts(df_phen_event$Emerge) <- contr.treatment(2, base = 1)
contrasts(df_phen_event$Trait) <- contr.treatment(2, base = 1)

m.emerge.peak <- lmer(Peak ~ 1 + Snowmelt*Emerge*Trait + Peak_Temp_new*Emerge*Trait + Year + (1 + Snowmelt + Peak_Temp_new | speciesplot), data=df_phen_event)
summary(m.emerge.peak)
```

```{r output=TRUE}
tab_model(m.emerge.peak, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
           terms = c("(Intercept)", "Snowmelt", "Year", "Peak_Temp_new", "Emerge2", "Trait2", "Snowmelt:Trait2"),
         pred.labels = c("Intercept", "Snowmelt", "Late active", "Surface Dwelling", "Temp", "Year", "Snowmelt:Surface dwelling"), title = "Peak ~ 1 + Snowmelt*Emerge + Temp*Emerge + Year + (1 + Snowmelt + Temp | SpeciesID:Plot)", dv.labels = "Early vs late active taxa", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
df_phen_event$Emerge <- as.factor(df_phen_event$Emerge)
df_phen_event$Trait <- as.factor(df_phen_event$Trait)
contrasts(df_phen_event$Emerge) <- contr.treatment(2, base = 1)
contrasts(df_phen_event$Trait) <- contr.treatment(2, base = 1)

m.emerge.dur <- lmer(Duration ~  1 + Snowmelt*Emerge*Trait + Onset_Temp_new*Emerge*Trait + Year + (1 + Snowmelt + Onset_Temp_new | speciesplot), data=df_phen_event)
summary(m.emerge.dur)
```

```{r output=TRUE}
tab_model(m.emerge.dur, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
           terms = c("(Intercept)", "Snowmelt", "Year", "Onset_Temp_new", "Emerge2", "Trait2", "Snowmelt:Trait2", "Emerge2:Trait2"),
         pred.labels = c("Intercept", "Snowmelt", "Late active", "Surface dwelling", "Temp", "Year", "Snowmelt:Surface dwelling", "Activity (late):Surface dwelling"), title = "Duration ~ 1 + Snowmelt*Emerge + Temp*Emerge + Year + (1 + Snowmelt + Temp | SpeciesID) + (1 + Snowmelt + Temp | SpeciesID:Plot)", dv.labels = "Taxa with different traits", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

In the table, only significant interactions are included.

```{r}
contrasts(df_phen_event$Emerge) <- contr.treatment(2, base = 1)
contrasts(df_phen_event$Trait) <- contr.treatment(2, base = 1)

m.emerge.peak <- lmer(Peak ~ 1 + Snowmelt*Emerge + Peak_Temp_new*Emerge + Year + (1 + Snowmelt + Peak_Temp_new | speciesplot), data=df_phen_event)
summary(m.emerge.peak)


m.trait.peak <- lmer(Peak ~ 1 + Snowmelt*Trait + Peak_Temp_new*Trait + Year + (1 + Snowmelt + Peak_Temp_new | speciesplot), data=df_phen_event)
summary(m.trait.peak)




df_phen_event$Emerge <- as.factor(df_phen_event$Emerge)
df_phen_event$Trait <- as.factor(df_phen_event$Trait)
contrasts(df_phen_event$Emerge) <- contr.treatment(2, base = 1)
contrasts(df_phen_event$Trait) <- contr.treatment(2, base = 1)

m.emerge.dur <- lmer(Duration ~  1 + Snowmelt*Emerge + Onset_Temp_new*Emerge + Year + (1 + Snowmelt + Onset_Temp_new | speciesplot), data=df_phen_event)
summary(m.emerge.dur)
```
