---
title: "Multiple_regression_arthropods"
author: "Hannah Sørine Gerlich"
format: html
editor: visual
message: false
warning: false
output: false
---

## Analyses on the effect of temperature and snowmelt on high-arctic arthropod phenology

This file contains all the code necessary to perform multiple regression modelling on the effect of temperature and snowmelt on each high-arctic arthropod phenology separately.

Initially, we load all the relevant packages.

```{r}
pkgs <- c("tidyverse", "readxl", "corrplot", "stats", "lme4", "lmerTest", "car",
          "sjPlot","MASS", "gtsummary")

vapply(pkgs, library, FUN.VALUE = logical(1L), character.only = TRUE,
       logical.return = TRUE)
```

Then read dataset with phenological metrics as well as air temperature estimates. The snowmelt data is in a separate csv file and we first have to bind the columns.

```{r}
df_phen_event <-
  read.csv(
    "Data/phenology_data/Pheno_air_temp.csv",
    sep = ",",
    stringsAsFactors = FALSE,
    header = TRUE
  )

dfsnowmelt_climatestation <-
  readxl::read_xlsx("Data/snowmelt_data/Snowmelt_climatestation.xlsx")

df_phen_event$Snowmelt <-
  dfsnowmelt_climatestation$DOY[match(paste0(df_phen_event$Year),
                                      paste0(dfsnowmelt_climatestation$Year))]
```

```{r}
df_phen_event %>%
  mutate(Order = case_when(
    SpeciesID == "Acari" ~ "Decomposer",
    SpeciesID == "ANMU" ~ "Mixed feeder",
    SpeciesID == "Aphidoidea" ~ "Herbivor",
    SpeciesID == "Chalcidoidea" ~ "Parasitoid",
    SpeciesID == "CHCE" ~ "Mixed feeder",
    SpeciesID == "Coccoidea" ~ "Herbivor",
    SpeciesID == "Collembola" ~ "Decomposer",
    SpeciesID == "Culicidae" ~ "Mixed feeder",
    SpeciesID == "Ichneumonidae" ~ "Parasitoid",
    SpeciesID == "Linyphiidae" ~ "Predator",
    SpeciesID == "Lycosidae" ~ "Predator",
    SpeciesID == "MYSC" ~ "Mixed feeder",
    SpeciesID == "Nymphalidae" ~ "Mixed feeder",
    SpeciesID == "Phoridae" ~ "Mixed feeder",
    SpeciesID == "Scathophagidae" ~ "Mixed feeder",
    SpeciesID == "Thomisidae" ~ "Predator")) -> df_phen_event

df_phen_event %>%
  mutate(Emerge = case_when(
    SpeciesID == "Acari" ~ "Early",
    SpeciesID == "ANMU" ~ "Late",
    SpeciesID == "Aphidoidea" ~ "Late",
    SpeciesID == "Chalcidoidea" ~ "Late",
    SpeciesID == "CHCE" ~ "Early",
    SpeciesID == "Coccoidea" ~ "Late",
    SpeciesID == "Collembola" ~ "Early",
    SpeciesID == "Culicidae" ~ "Late",
    SpeciesID == "Ichneumonidae" ~ "Late",
    SpeciesID == "Linyphiidae" ~ "Early",
    SpeciesID == "Lycosidae" ~ "Early",
    SpeciesID == "MYSC" ~ "Early",
    SpeciesID == "Nymphalidae" ~ "Late",
    SpeciesID == "Phoridae" ~ "Late",
    SpeciesID == "Scathophagidae" ~ "Late",
    SpeciesID == "Thomisidae" ~ "Early")) -> df_phen_event

df_phen_event %>%
  mutate(Trait = case_when(
    SpeciesID == "Acari" ~ "Surface",
    SpeciesID == "ANMU" ~ "Flying",
    SpeciesID == "Aphidoidea" ~ "Surface",
    SpeciesID == "Chalcidoidea" ~ "Flying",
    SpeciesID == "CHCE" ~ "Flying",
    SpeciesID == "Coccoidea" ~ "Surface",
    SpeciesID == "Collembola" ~ "Surface",
    SpeciesID == "Culicidae" ~ "Flying",
    SpeciesID == "Ichneumonidae" ~ "Flying",
    SpeciesID == "Linyphiidae" ~ "Surface",
    SpeciesID == "Lycosidae" ~ "Surface",
    SpeciesID == "MYSC" ~ "Flying",
    SpeciesID == "Nymphalidae" ~ "Flying",
    SpeciesID == "Phoridae" ~ "Flying",
    SpeciesID == "Scathophagidae" ~ "Flying",
    SpeciesID == "Thomisidae" ~ "Surface")) -> df_phen_event
```

We find strong collinearity between temperature and snowmelt for families in the snowbed plot (Art6) and Nymphalidae in one mesic heath plot (Art3). We also do not consider phenology observations for spiders (predators) and decomposers in the window trap (Plot 1) reliable. Therefore, we have to remove these observations from the data.

```{r}
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art1" &
                    (df_phen_event$Order == "Decomposer")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art1" &
                    (df_phen_event$Order == "Predator")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art3" &
                    (df_phen_event$SpeciesID == "Nymphalidae")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$SpeciesID == "Lygaeidae"), ]

df_phen_event <- subset(df_phen_event,!Plot == "Art6")
```

Assign habitat type.

```{r}
df_phen_event %>%
  mutate(
    Habitat = case_when(
      Plot == "Art1" ~ "Pond",
      Plot == "Art2" ~ "Wet fen",
      Plot == "Art3" ~ "Mesic heath",
      Plot == "Art4" ~ "Mesic heath",
      Plot == "Art5" ~ "Arid heath",
      Plot == "Art7" ~ "Arid heath"
    )
  ) -> df_phen_event
```

Remove taxa in the dataset for which we do not have enough data.

```{r}
df_phen_event$Order[is.na(df_phen_event$Order)] <- "NA"
df_phen_event <- subset(df_phen_event, Order != "NA")
```

Multiple regression is then performed on each individual taxa. The best model is chosen based on AIC model comparison and anova results. This is also to test which climate variable is most important in determining each individual arthropod taxa phenology.

Start with decomposers (Acari and Collembola).

## Acari

```{r}
df_Acari <- subset(df_phen_event,SpeciesID=="Acari")
df_Acari%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Acari
```

We are also interested in comparing responses among habitats. To do this with Plot as a categorical variable, we use costumized contrasts to specify which habitats we want to compare.

```{r}
unique(df_Acari$Plot)
#Plot 2 (wet fen), Plot 3, 4 (mesic heath) and Plot 5, 7 (arid heath)

df_Acari$Plot <- as.factor(df_Acari$Plot)


#mycontrasts = contrasts(df_CHCE$Plot)

#S <- cbind(1, contrasts(df_CHCE$Plot))
#L <- ginv(S)
#L %>% fractions()
H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
L <- rbind(H1, H2) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Acari$Plot) <- S
contrasts(df_Acari$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Acari)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Acari)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Acari)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Acari)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Acari)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Acari)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30+Plot+Year, df_Acari)

ancova_acari <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_acari

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

The results for Acari show no effect of snowmelt or temperature on peak phenology. There are differences in phenological responses to climate variables among habitats, though.

```{r output=TRUE}
tbl_regression(multiple_lm1)
```

```{r}
H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Acari$Plot) <- S
contrasts(df_Acari$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Acari)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Acari)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Acari)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Acari)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Acari)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Acari)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30+Plot+Year, df_Acari)

ancova_acari <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_acari

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

## Collembola

```{r}
df_Collembola <- subset(df_phen_event,SpeciesID=="Collembola")
df_Collembola%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Collembola

unique(df_Collembola$Plot)
#Plot 2 (wet fen), Plot 3, 4 (mesic heath) and Plot 5, 7 (arid heath)

df_Collembola$Plot <- as.factor(df_Collembola$Plot)

H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
L <- rbind(H1, H2) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Collembola$Plot) <- S
contrasts(df_Collembola$Plot)

```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Collembola)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Collembola)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Collembola)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Collembola)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Collembola)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Collembola)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30+Plot+Year, df_Collembola)

ancova_Collembola <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_Collembola

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

The best model is with plot and temperature as well as snowmelt interactions.

```{r output=TRUE}
tbl_regression(multiple_lm4)
```

```{r}
H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Collembola$Plot) <- S
contrasts(df_Collembola$Plot)

```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Collembola)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Collembola)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Collembola)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Collembola)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Collembola)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Collembola)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30+Plot+Year, df_Collembola)


ancova_Collembola <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_Collembola

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

```{r output=TRUE}
tbl_regression(multiple_lm4)
```

The phenological responses to climate variables differ between mesic heath and arid heath habitats.

# Herbivores

## Aphidoidea

We only have phenological metrics for plot 7 and therefore the sutomized constrasting is not relevant here.

```{r}
df_APHI <- subset(df_phen_event,SpeciesID=="Aphidoidea")
df_APHI%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_APHI

multiple_lm0a <- lm(Peak ~ Snowmelt+Year, df_APHI)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Year, df_APHI)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Year, df_APHI)
multiple_lm2 <- lm(Peak ~ Snowmelt*Peak_Temp30+Year, df_APHI)

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2)
```

```{r}
tbl_regression(multiple_lm1)
```

The effect of climate variables on aphidoidea phenology is not strong.

## Coccoidea

```{r}
df_Cocco <- subset(df_phen_event,SpeciesID=="Coccoidea")
df_Cocco%>%
  subset(!is.na(Onset_Temp30)&!is.na(Onset)) -> df_Cocco

unique(df_Cocco$Plot)
#Plot 3, 4 (mesic heath) and plot 5, 7 (arid heath)

df_Cocco$Plot <- as.factor(df_Cocco$Plot)


H1 = c(1/2, 1/2, -1/2, -1/2)
L <- rbind(H1) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Cocco$Plot) <- S
contrasts(df_Cocco$Plot)

```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Cocco)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Cocco)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Cocco)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Cocco)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Cocco)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Cocco)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30+Plot+Year, df_Cocco)

ancova_Cocco <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_Cocco

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

```{r output=TRUE}
tbl_regression(multiple_lm2)
```

Coccoidea exhibits strong earlier peak phenology in response to snowmelt.

# Parasitoids

## Chalcidoidea

```{r}
df_Chal <- subset(df_phen_event,SpeciesID=="Chalcidoidea")
df_Chal%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Chal

unique(df_Chal$Plot)
#Plot 3, 4 (mesic heath) and plot 5, 7 (arid heath)

df_Chal$Plot <- as.factor(df_Chal$Plot)

H1 = c(1/2, 1/2, -1/2, -1/2)
L <- rbind(H1) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Chal$Plot) <- S
contrasts(df_Chal$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Chal)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Chal)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Chal)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Chal)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Chal)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Chal)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30+Plot+Year, df_Chal)

ancova_Chal <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_Chal

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

```{r}
tbl_regression(multiple_lm1)
```

We find no differences among habitats. Snowmelt has a large effect on peak phenology.

## Ichneumonidae

We need to consider the random effect of trap type for Ichneumonidae, because plot 1 is a window trap and this type of trap is substantially different from pitfall traps in the other plots.

```{r}
df_Ich <- subset(df_phen_event,SpeciesID=="Ichneumonidae")
df_Ich%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Ich

df_Ich%>%
  mutate(Trap = case_when(
    endsWith(Plot, "1") ~ "Window",
    endsWith(Plot, "2") ~ "Pitfall",
    endsWith(Plot, "3") ~ "Pitfall",
    endsWith(Plot, "4") ~ "Pitfall",
    endsWith(Plot, "5") ~ "Pitfall",
    endsWith(Plot, "7") ~ "Pitfall"
    )) -> df_Ich

df_Ich$Trap <- as.factor(df_Ich$Trap) 
df_Ich$Plot <- as.factor(df_Ich$Plot)


H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H1, H2, H4) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Ich$Plot) <- S
contrasts(df_Ich$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Ich)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Ich)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Ich)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Ich)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Ich)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Ich)

multiple_trap_lm0a <- lmer(Peak ~ Snowmelt+Plot+Year+(1|Plot:Trap), df_Ich)
multiple_trap_lm0b <- lmer(Peak ~ Peak_Temp30+Plot+Year+(1|Plot:Trap), df_Ich)
multiple_trap_lm1 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+(1|Plot:Trap), df_Ich)
multiple_trap_lm2 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+(1|Plot:Trap), df_Ich)
multiple_trap_lm3 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot+(1|Plot:Trap), df_Ich)
multiple_trap_lm4 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot+(1|Plot:Trap), df_Ich)

ancova_Ich <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
ancova_Ich

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)

ancova_Ich <- anova(multiple_trap_lm0a, multiple_trap_lm0b, multiple_trap_lm1, multiple_trap_lm2, multiple_trap_lm3, multiple_trap_lm4)
ancova_Ich

AIC(multiple_trap_lm0a, multiple_trap_lm0b, multiple_trap_lm1, multiple_trap_lm2, multiple_trap_lm3, multiple_trap_lm4)
```

```{r output=TRUE}
tbl_regression(multiple_lm1)
```

We find no differences among habitats. Snowmelt and temperature both have large effects on peak phenology.

# Mixed feeders

## Chironomidae

We need to consider the random effect of trap type for Chironomoidae, because plot 1 is a window trap and this type of trap is substantially different from pitfall traps in the other plots.

```{r}
df_CHCE <- subset(df_phen_event,SpeciesID=="CHCE")
df_CHCE%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_CHCE

unique(df_CHCE$Plot)
#Plot 1 (pond), plot 2, (wet fen), plot 3, 4 (mesic heath) and plot 5, 7 (arid heath)

df_CHCE%>%
  mutate(Trap = case_when(
    endsWith(Plot, "1") ~ "Window",
    endsWith(Plot, "2") ~ "Pitfall",
    endsWith(Plot, "3") ~ "Pitfall",
    endsWith(Plot, "4") ~ "Pitfall",
    endsWith(Plot, "5") ~ "Pitfall",
    endsWith(Plot, "7") ~ "Pitfall"
    )) -> df_CHCE

df_CHCE$Trap <- as.factor(df_CHCE$Trap) 
df_CHCE$Plot <- as.factor(df_CHCE$Plot)

#mycontrasts = contrasts(df_CHCE$Plot)

#S <- cbind(1, contrasts(df_CHCE$Plot))
#L <- ginv(S)
#L %>% fractions()
H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H1, H2, H4) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_CHCE$Plot) <- S
contrasts(df_CHCE$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_CHCE)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_CHCE)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_CHCE)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_CHCE)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_CHCE)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_CHCE)

multiple_trap_lm0a <- lmer(Peak ~ Snowmelt+Plot+Year+(1|Plot:Trap), df_CHCE)
multiple_trap_lm0b <- lmer(Peak ~ Peak_Temp30+Plot+Year+(1|Plot:Trap), df_CHCE)
multiple_trap_lm1 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+(1|Plot:Trap), df_CHCE)
multiple_trap_lm2 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+(1|Plot:Trap), df_CHCE)
multiple_trap_lm3 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot+(1|Plot:Trap), df_CHCE)
multiple_trap_lm4 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot+(1|Plot:Trap), df_CHCE)

ancova_CHCE <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
ancova_CHCE

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)

ancova_CHCE <- anova(multiple_trap_lm0a, multiple_trap_lm0b, multiple_trap_lm1, multiple_trap_lm2, multiple_trap_lm3, multiple_trap_lm4)
ancova_CHCE

AIC(multiple_trap_lm0a, multiple_trap_lm0b, multiple_trap_lm1, multiple_trap_lm2, multiple_trap_lm3, multiple_trap_lm4)
```

```{r output=TRUE}
tbl_regression(multiple_trap_lm1)
```

```{r}
H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_CHCE$Plot) <- S
contrasts(df_CHCE$Plot)
```

```{r}
multiple_trap_lm0a <- lmer(Peak ~ Snowmelt+Plot+Year+(1|Plot:Trap), df_CHCE)
multiple_trap_lm0b <- lmer(Peak ~ Peak_Temp30+Plot+Year+(1|Plot:Trap), df_CHCE)
multiple_trap_lm1 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+(1|Plot:Trap), df_CHCE)
multiple_trap_lm2 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+(1|Plot:Trap), df_CHCE)
multiple_trap_lm3 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot+(1|Plot:Trap), df_CHCE)
multiple_trap_lm4 <- lmer(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot+(1|Plot:Trap), df_CHCE)

ancova_CHCE <- anova(multiple_trap_lm0a, multiple_trap_lm0b, multiple_trap_lm1, multiple_trap_lm2, multiple_trap_lm3, multiple_trap_lm4)
ancova_CHCE

AIC(multiple_trap_lm0a, multiple_trap_lm0b, multiple_trap_lm1, multiple_trap_lm2, multiple_trap_lm3, multiple_trap_lm4)
```

```{r output=TRUE}
tbl_regression(multiple_trap_lm1)
```

## Muscidae

There was no substantial variance explained from the random effect of trap type for Muscidae. Therefore, we do not consider it relevant to add this covariate in the model.

```{r}
df_ANMU <- subset(df_phen_event,SpeciesID=="ANMU")
df_ANMU%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_ANMU

unique(df_ANMU$Plot)

#Plot 1 (pond), plot 2, (wet fen), plot 3, 4 (mesic heath) and plot 5, 7 (arid heath)


df_ANMU$Plot <- as.factor(df_ANMU$Plot)

H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H1, H2, H4) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_ANMU$Plot) <- S
contrasts(df_ANMU$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_ANMU)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_ANMU)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_ANMU)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_ANMU)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_ANMU)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_ANMU)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30*Plot+Year, df_ANMU)

ancova_ANMU <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_ANMU

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

```{r output=TRUE}
tbl_regression(multiple_lm2)
```

```{r}
H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_ANMU$Plot) <- S
contrasts(df_ANMU$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_ANMU)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_ANMU)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_ANMU)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_ANMU)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_ANMU)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_ANMU)

ancova_ANMU <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
ancova_ANMU

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
```

```{r output=TRUE}
tbl_regression(multiple_lm2)
```

## Culicidae

We only have phenological metrics for plot 1 and therefore the sutomized constrasting is not relevant here.

```{r}
df_culi <- subset(df_phen_event,SpeciesID=="Culicidae")
df_culi%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_culi

multiple_lm0a <- lm(Peak ~ Snowmelt+Year, df_culi)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Year, df_culi)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Year, df_culi)

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1)
```

```{r}
tbl_regression(multiple_lm1)
```

Timing of snowmelt has a significant effect on earlier peak activity for Culicidae. The temperature variable was not significant, but still explained a substantial amount of variation in the model (model with both variables was preferred).

## Sciaridae

There was no substantial variance explained from the random effect of trap type for Sciaridae. Therefore, we do not consider it relevant to add this covariate in the model.

```{r}

df_MYSC <- subset(df_phen_event,SpeciesID=="MYSC")

df_MYSC%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_MYSC

unique(df_MYSC$Plot)

df_MYSC$Plot <- as.factor(df_MYSC$Plot)


#mycontrasts = contrasts(df_ANMU$Plot)

#S <- cbind(1, contrasts(df_ANMU$Plot))
#L <- ginv(S)
#L %>% fractions()
H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H1, H2, H4) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_MYSC$Plot) <- S
contrasts(df_MYSC$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_MYSC)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_MYSC)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_MYSC)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_MYSC)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_MYSC)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_MYSC)

ancova_MYSC <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
ancova_MYSC

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
```

```{r}
tbl_regression(multiple_lm1)
```

```{r}
H1 = c(0, 0, 1/2, 1/2, -1/2, -1/2)
H2 = c(0, -1, 1/2, 1/2, 0, 0)
H3 = c(0, 1, 0, 0, -1/2, -1/2)
H4 = c(1, -1/5, -1/5, -1/5, -1/5, -1/5)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_MYSC$Plot) <- S
contrasts(df_MYSC$Plot)
```

```{r}
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_MYSC)
```

```{r}
tbl_regression(multiple_lm1)
```

Sciaridae responded strongly to timing of snowmelt and temperature. A simple model with no interactions was preferred, however, we did find differences between early and late snow melting habitats.

## Phoridae

```{r}
df_Phori <- subset(df_phen_event,SpeciesID=="Phoridae")
df_Phori%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Phori

unique(df_Phori$Plot)
#Plot 3, 4 (mesic heath) and plot 5, 7 (arid heath)

df_Phori$Plot <- as.factor(df_Phori$Plot)


#mycontrasts = contrasts(df_ANMU$Plot)

#S <- cbind(1, contrasts(df_ANMU$Plot))
#L <- ginv(S)
#L %>% fractions()
H1 = c(1/2, 1/2, -1/2, -1/2)
L <- rbind(H1) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Phori$Plot) <- S
contrasts(df_Phori$Plot)

```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Phori)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Phori)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Phori)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Phori)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Phori)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Phori)

ancova_Phori <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
ancova_Phori

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4)
```

```{r}
tbl_regression(multiple_lm1)
```

There is a strong effect of timing of snowmelt on earlier phoridae peak phenology, and there is also a negative effect of temperature, but the estimate is not significant.

## Nymphalidae

```{r}
df_Nymph <- subset(df_phen_event,SpeciesID=="Nymphalidae")
df_Nymph%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Nymph

multiple_lm0a <- lm(Peak ~ Snowmelt+Year, df_Nymph)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Year, df_Nymph)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Year, df_Nymph)

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1)
```

```{r}
tbl_regression(multiple_lm1)
```

# Predators

## Linyphiidae

```{r}
df_Liny <- subset(df_phen_event,SpeciesID=="Linyphiidae")
df_Liny%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Liny

df_Liny$Plot <- as.factor(df_Liny$Plot)

H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
#H4 = c(-1/5, -1/5, -1/5, -1/5, 1, -1/5)
L <- rbind(H1, H2) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Liny$Plot) <- S
contrasts(df_Liny$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Liny)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Liny)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Liny)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Liny)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Liny)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Liny)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30*Plot+Year, df_Liny)

ancova_Liny <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_Liny

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

```{r}
H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
#H4 = c(-1/5, -1/5, -1/5, -1/5, 1, -1/5)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Liny$Plot) <- S
contrasts(df_Liny$Plot)
```

```{r}
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Liny)
```

```{r}
tbl_regression(multiple_lm1)
```

Linyphiidae is the only taxa to show a delayed peak activity in response to earlier snowmelt.

## Lycosidae

```{r}
df_Lyco <- subset(df_phen_event,SpeciesID=="Lycosidae")
df_Lyco%>%
  subset(!is.na(Onset_Temp30)&!is.na(Onset)) -> df_Lyco

df_Lyco$Plot <- as.factor(df_Lyco$Plot)

H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
L <- rbind(H1, H2) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Lyco$Plot) <- S
contrasts(df_Lyco$Plot)
```

```{r}
multiple_lm0a <- lm(Peak ~ Snowmelt+Plot+Year, df_Lyco)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Plot+Year, df_Lyco)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Lyco)
multiple_lm2 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot, df_Lyco)
multiple_lm3 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Peak_Temp30:Plot, df_Lyco)
multiple_lm4 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year+Snowmelt:Plot+Peak_Temp30:Plot, df_Lyco)
multiple_lm5 <- lm(Peak ~ Snowmelt*Peak_Temp30*Plot+Year, df_Lyco)

ancova_Lyco <- anova(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
ancova_Lyco

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1, multiple_lm2, multiple_lm3, multiple_lm4, multiple_lm5)
```

```{r}
H1 = c(0, 1/2, 1/2, -1/2, -1/2)
H2 = c(-1, 1/2, 1/2, 0, 0)
H3 = c(1, 0, 0, -1/2, -1/2)
L <- rbind(H3) # bind the rows together
L %>% fractions()
S <- ginv(L) # compute the inverse
S 

contrasts(df_Lyco$Plot) <- S
contrasts(df_Lyco$Plot)
```

```{r}
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Plot+Year, df_Lyco)
```

```{r}
tbl_regression(multiple_lm1)
```

Lycosidae responds strongly to earlier snowmelt and increased temperatures in peak phenology.

## Thomisidae

```{r}
df_Thomi <- subset(df_phen_event,SpeciesID=="Thomisidae")
df_Thomi%>%
  subset(!is.na(Peak_Temp30)&!is.na(Peak)) -> df_Thomi

multiple_lm0a <- lm(Peak ~ Snowmelt+Year, df_Thomi)
multiple_lm0b <- lm(Peak ~ Peak_Temp30+Year, df_Thomi)
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Year, df_Thomi)

AIC(multiple_lm0a, multiple_lm0b, multiple_lm1)
```

```{r}
multiple_lm1 <- lm(Peak ~ Snowmelt+Peak_Temp30+Year, df_Thomi)
```

```{r}
tbl_regression(multiple_lm1)
```

No strong effect of climate variables on peak phenology.
