---
title: "Pheno_responses_climate"
author: "Hannah Sørine Gerlich"
format: html
editor: visual
---

# Analyses on the effect of temperature and snowmelt on high-arctic arthropod phenology

This file contains all the code necessary to perform linear mixed modelling on the effect of temperature and snowmelt on high-arctic arthropod phenology.

Initially, we load all the relevant packages.

```{r}
#| echo: false
#| message: false
#| warning: false

pkgs <- c("tidyverse", "readxl", "corrplot", "stats", "lme4", "lmerTest", "car",
          "sjPlot")

vapply(pkgs, library, FUN.VALUE = logical(1L), character.only = TRUE,
       logical.return = TRUE)
```

Then read dataset with phenological metrics as well as air temperature estimates. The snowmelt data is in a separate csv file and we first have to bind the columns.

```{r}
#| message: false
#| warning: false

df_phen_event <-
  read.csv(
    "Data/phenology_data/Pheno_air_temp.csv",
    sep = ",",
    stringsAsFactors = FALSE,
    header = TRUE
  )

dfsnowmelt_climatestation <-
  readxl::read_xlsx("Data/snowmelt_data/Snowmelt_climatestation.xlsx")

df_phen_event$Snowmelt <-
  dfsnowmelt_climatestation$DOY[match(paste0(df_phen_event$Year),
                                      paste0(dfsnowmelt_climatestation$Year))]
```

```{r}
#| echo: false
#| message: false
#| warning: false

df_phen_event %>%
  mutate(Order = case_when(
    SpeciesID == "Acari" ~ "Decomposer",
    SpeciesID == "ANMU" ~ "Mixed feeder",
    SpeciesID == "Aphidoidea" ~ "Herbivor",
    SpeciesID == "Chalcidoidea" ~ "Parasitoid",
    SpeciesID == "CHCE" ~ "Mixed feeder",
    SpeciesID == "Coccoidea" ~ "Herbivor",
    SpeciesID == "Collembola" ~ "Decomposer",
    SpeciesID == "Culicidae" ~ "Mixed feeder",
    SpeciesID == "Ichneumonidae" ~ "Parasitoid",
    SpeciesID == "Linyphiidae" ~ "Predator",
    SpeciesID == "Lycosidae" ~ "Predator",
    SpeciesID == "MYSC" ~ "Mixed feeder",
    SpeciesID == "Nymphalidae" ~ "Mixed feeder",
    SpeciesID == "Phoridae" ~ "Mixed feeder",
    SpeciesID == "Thomisidae" ~ "Predator")) -> df_phen_event

df_phen_event %>%
  mutate(Emerge = case_when(
    SpeciesID == "Acari" ~ "Early",
    SpeciesID == "ANMU" ~ "Late",
    SpeciesID == "Aphidoidea" ~ "Late",
    SpeciesID == "Chalcidoidea" ~ "Late",
    SpeciesID == "CHCE" ~ "Early",
    SpeciesID == "Coccoidea" ~ "Late",
    SpeciesID == "Collembola" ~ "Early",
    SpeciesID == "Culicidae" ~ "Late",
    SpeciesID == "Ichneumonidae" ~ "Late",
    SpeciesID == "Linyphiidae" ~ "Early",
    SpeciesID == "Lycosidae" ~ "Early",
    SpeciesID == "MYSC" ~ "Early",
    SpeciesID == "Nymphalidae" ~ "Late",
    SpeciesID == "Phoridae" ~ "Late",
    SpeciesID == "Scathophagidae" ~ "Late",
    SpeciesID == "Thomisidae" ~ "Early")) -> df_phen_event

df_phen_event %>%
  mutate(Trait = case_when(
    SpeciesID == "Acari" ~ "Surface",
    SpeciesID == "ANMU" ~ "Flying",
    SpeciesID == "Aphidoidea" ~ "Surface",
    SpeciesID == "Chalcidoidea" ~ "Flying",
    SpeciesID == "CHCE" ~ "Flying",
    SpeciesID == "Coccoidea" ~ "Surface",
    SpeciesID == "Collembola" ~ "Surface",
    SpeciesID == "Culicidae" ~ "Flying",
    SpeciesID == "Ichneumonidae" ~ "Flying",
    SpeciesID == "Linyphiidae" ~ "Surface",
    SpeciesID == "Lycosidae" ~ "Surface",
    SpeciesID == "MYSC" ~ "Flying",
    SpeciesID == "Nymphalidae" ~ "Flying",
    SpeciesID == "Phoridae" ~ "Flying",
    SpeciesID == "Scathophagidae" ~ "Flying",
    SpeciesID == "Thomisidae" ~ "Surface")) -> df_phen_event
```

We find strong collinearity between temperature and snowmelt for families in the snowbed plot (Art6) and Nymphalidae in one mesic heath plot (Art3). Therefore, we have to remove these observations from the data.

```{r}
#| message: false
#| warning: false

df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art1" &
                    (df_phen_event$Order == "Decomposer")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art6" &
                    (df_phen_event$Order == "Decomposer")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art6" &
                    (df_phen_event$Order == "Predator")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art1" &
                    (df_phen_event$Order == "Predator")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art6" &
                    (df_phen_event$Order == "Parasitoid")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art6" &
                    (df_phen_event$Order == "Mixed feeder")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$Plot == "Art3" &
                    (df_phen_event$SpeciesID == "Nymphalidae")), ]
df_phen_event <-
  df_phen_event[!(df_phen_event$SpeciesID == "Lygaeidae"), ]

df_phen_event <- subset(df_phen_event,!Plot == "Art6")
```

```{r}
#| echo: false
#| message: false
#| warning: false

df_phen_event %>%
  mutate(
    Habitat = case_when(
      Plot == "Art1" ~ "Pond",
      Plot == "Art2" ~ "Wet fen",
      Plot == "Art3" ~ "Mesic heath",
      Plot == "Art4" ~ "Mesic heath",
      Plot == "Art5" ~ "Arid heath",
      Plot == "Art7" ~ "Arid heath"
    )
  ) -> df_phen_event
```

```{r}
#| echo: false
#| message: false
#| warning: false

df_phen_event$Order[is.na(df_phen_event$Order)] <- "NA"
df_phen_event <- subset(df_phen_event, Order != "NA")
```

Because the temperature variable is specific for each taxon-by-plot combination, we need to account for within subject variation. We also have to add year as a covariate to detrend the model.

```{r}
#| echo: false
#| message: false
#| warning: false

df_phen_event$speciesplot <-
  as.factor(paste(df_phen_event$SpeciesID, df_phen_event$Plot))

df_phen_event$Peak_Temp_new <-
  df_phen_event$Peak_Temp30 - tapply(df_phen_event$Peak_Temp30, df_phen_event$speciesplot, mean)[df_phen_event$speciesplot]
df_phen_event$Onset_Temp_new <-
  df_phen_event$Onset_Temp30 - tapply(df_phen_event$Onset_Temp30, df_phen_event$speciesplot, mean)[df_phen_event$speciesplot]
df_phen_event$End_Temp_new <-
  df_phen_event$End_Temp30 - tapply(df_phen_event$End_Temp30, df_phen_event$speciesplot, mean)[df_phen_event$speciesplot]
```

Because data for some taxa in some plots are sparse, we scale the predictors so that the model can converge.

```{r}
#| message: false
#| warning: false

df_phen_event$peaktemp_scaled <- scale(df_phen_event$Peak_Temp_new, center = F)
df_phen_event$snow_scaled <- scale(df_phen_event$Snowmelt, center = F)
df_phen_event$peak_scaled <- scale(df_phen_event$Peak, center = F)

colnames(df_phen_event)[24] ="peaktemp_scaled"
colnames(df_phen_event)[25] ="snow_scaled"
colnames(df_phen_event)[26] ="peak_scaled"

df_phen_event$onsettemp_scaled <- scale(df_phen_event$Onset_Temp_new, center = F)
df_phen_event$duration_scaled <- scale(df_phen_event$Duration, center = F)
df_phen_event$onset_scaled <- scale(df_phen_event$Onset, center = F)

colnames(df_phen_event)[27] ="onsettemp_scaled"
colnames(df_phen_event)[28] ="duration_scaled"
colnames(df_phen_event)[29] ="onset_scaled"

df_phen_event$endtemp_scaled <- scale(df_phen_event$End_Temp_new, center = F)
df_phen_event$end_scaled <- scale(df_phen_event$End, center = F)

colnames(df_phen_event)[30] ="endtemp_scaled"
colnames(df_phen_event)[31] ="end_scaled"
```

The linear mixed model now contains scaled climate variables with relevant random slope and intercept effects of taxa and plot on each climate variable. Also, sampling variance (speciesplot) has been added.

```{r}
#| echo: false
#| message: false
#| warning: false
#| output: false

m.peak <-
  lmer(
    Peak ~ 1 + snow_scaled + peaktemp_scaled + Year + 
      (1 | Plot) +
      (1 + snow_scaled + peaktemp_scaled | speciesplot) +
      (1 + snow_scaled + peaktemp_scaled | SpeciesID) +
      (1 + snow_scaled + peaktemp_scaled | Habitat),
    data = df_phen_event
  )

summary(m.peak)

peak_coefs <- summary(m.peak)$coefficients[2,1:ncol(summary(m.peak)$coefficients)]
peak_coefs <- data.frame(summary(m.peak)$coefficients[c(2:3),c(1:3)])
peak_coefs$variable <- c("Snowmelt", "Temperature")
peak_coefs$pheno <- "Peak"

```

Create table from effect of climate variables on community peak phenology.

**Table S5.1** Effect sizes of snowmelt timing and temperature on arthropod community peak activity. Random intercepts and slopes of taxa and habitat as well as sampling variance have been included, but results are not shown in table.

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.peak, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE, show.re.var = FALSE,
         title = "Peak ~ Snowmelt + Temp + Year + (1 + Temp + snow | Taxa) + (1 + Temp + snow | Plot)", terms = c("(Intercept)", "snow_scaled", "peaktemp_scaled", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.duration <-
  lmer(
    Duration ~ 1 + snow_scaled + onsettemp_scaled + Year + 
      (1 | Plot) +
      (1 + snow_scaled + onsettemp_scaled | speciesplot) +
      (1 + snow_scaled + onsettemp_scaled | SpeciesID) +
      (1 + snow_scaled + onsettemp_scaled | Habitat),
    data = df_phen_event
  )

summary(m.duration)

duration_coefs <- summary(m.duration)$coefficients[2,1:ncol(summary(m.duration)$coefficients)]
duration_coefs <- data.frame(summary(m.duration)$coefficients[c(2:3),c(1:3)])
duration_coefs$variable <- c("Snowmelt", "Temperature")
duration_coefs$pheno <- "Duration"
```

**Table S5.2** Effect sizes of snowmelt timing and temperature on arthropod community duration of activity. Random intercepts and slopes of taxa and habitat as well as sampling variance have been included, but results are not shown in table.

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.duration, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE, show.re.var = FALSE,
         title = "Duration ~ Snowmelt + Temp + Year + (1 + Temp + snow | Taxa) + (1 + Temp + snow | Plot)", terms = c("(Intercept)", "snow_scaled", "onsettemp_scaled", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.onset <-
  lmer(
    Onset ~ 1 + snow_scaled + onsettemp_scaled + Year + 
      (1 | Plot) +
      (1 + snow_scaled + onsettemp_scaled | speciesplot) +
      (1 + snow_scaled + onsettemp_scaled | SpeciesID) +
      (1 + snow_scaled + onsettemp_scaled | Habitat),
    data = df_phen_event
  )

summary(m.onset)

onset_coefs <- summary(m.onset)$coefficients[2,1:ncol(summary(m.onset)$coefficients)]
onset_coefs <- data.frame(summary(m.onset)$coefficients[c(2:3),c(1:3)])
onset_coefs$variable <- c("Snowmelt", "Temperature")
onset_coefs$pheno <- "Onset"
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.end <-
  lmer(
    End ~ 1 + snow_scaled + endtemp_scaled + Year + 
      (1 | Plot) +
      (1 + snow_scaled + endtemp_scaled | speciesplot) +
      (1 + snow_scaled + endtemp_scaled | SpeciesID) +
      (1 + snow_scaled + endtemp_scaled | Habitat),
    data = df_phen_event
  )

summary(m.end)

end_coefs <- summary(m.end)$coefficients[2,1:ncol(summary(m.end)$coefficients)]
end_coefs <- data.frame(summary(m.end)$coefficients[c(2:3),c(1:3)])
end_coefs$variable <- c("Snowmelt", "Temperature")
end_coefs$pheno <- "End"
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

df_combine <- bind_rows(onset_coefs,peak_coefs,end_coefs,duration_coefs)
```

```{r}
#| message: false
#| warning: false
#| echo: false

df_combine$pheno <- factor(df_combine$pheno,                 # Relevel group factor
                         levels = c("Onset", "Peak", "End", "Duration"))

ggplot(df_combine, aes(pheno, Estimate, fill = pheno)) +
  xlab("") +
  ylab("Standardized effect size") +
  coord_cartesian(ylim=c(-25, 65)) +
  geom_errorbar(
    aes(
      ymin = Estimate - Std..Error,
      ymax = Estimate + Std..Error,
      color = pheno
    ),
    alpha = 0.8,
    width = 0.1,
    linewidth = 1
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 3,
             shape = 21,
             color = "black") +
  facet_grid(. ~ variable) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.spacing = unit(.08, "lines"),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      size = 0.5),
      axis.text.y = element_text(hjust = 0.5, size = 12),
      axis.text.x = element_text(angle = 35, vjust = 0.5, size = 12),
      axis.title.y = element_text(vjust = 2, size = 14)
      #plot.margin = margin(10,4,10,4,"cm")
    )

```

**Figure S5.1** The relative effect of snowmelt timing and temperature across all arthropod phenological events (onset, peak, end and duration of activity). Standardized effect sizes are defined as the slope coefficients derived from scaled and centred explanatory variables. Timing of snowmelt is the primary driver of arthropod phenology, with earlier onset, peak and end of phenology as the snow melts earlier. The duration of activity extends with earlier snowmelt timing. The effect of temperature on phenology is substantially smaller, but temperature is significantly correlated with phenology such that arthropod phenology starts earlier with warming.

```{r}
#| message: false
#| warning: false
#| echo: false

df_combine$pheno <- factor(df_combine$pheno,                 # Relevel group factor
                         levels = c("Duration", "End", "Peak", "Onset"))

ggplot(df_combine, aes(Estimate, pheno, fill = pheno)) +
     xlab("Standardized effect size") +
     ylab("") +
     coord_cartesian(xlim=c(-25, 65)) +
     geom_errorbar(
         aes(
             xmin = Estimate - Std..Error,
             xmax = Estimate + Std..Error,
             color = pheno
         ),
         alpha = 0.8,
         width = 0.1,
         linewidth = 1
     ) +
     geom_vline(xintercept = 0, linetype = "dashed") +
     geom_point(size = 3,
                shape = 21,
                color = "black") +
     facet_grid(. ~ variable) +
     scale_fill_viridis_d(option = "plasma") +
     scale_color_viridis_d(option = "plasma") +
     theme_bw() +
     theme(
         panel.background = element_rect(fill = "white"),
         panel.spacing = unit(.08, "lines"),
         panel.border = element_rect(
             color = "black",
             fill = NA,
             size = 0.5),
         axis.text.y = element_text(hjust = 0.5, size = 12),
         axis.text.x = element_text(vjust = 0.5, size = 12),
         axis.title.x = element_text(vjust = -1, size = 14),
         legend.position = "none"
         #plot.margin = margin(10,2,10,2,"cm")
     )
```

```{r}
#| message: false
#| warning: false
#| output: false
#| echo: false

library(ggpubr)

#Visualize community trends
peak <- ggplot(df_phen_event, aes(x = Snowmelt, y = Peak, color = SpeciesID)) +
     #facet_wrap(~ Habitat, nrow=2) +   # a panel for each mountain range
     geom_point(alpha = 0.5) +
     ylab("Peak activity (Day of Year)") +
     xlab("Snowmelt timing (Day of Year)") +
     theme_classic() +
     scale_color_viridis_d(option = "plasma") +
     geom_smooth(se = FALSE, method = "lm", linewidth = 2) +
     theme(legend.position = "none",
           axis.title.y = element_text(vjust = 2, size = 14),
           axis.title.x = element_text(vjust = 0.8, size = 14),
           axis.text.y = element_text(hjust = 0.5, size = 12),
           axis.text.x = element_text(angle = 35, vjust = 0.5, size = 12),
           plot.margin = margin(10,1,10,0.5, "cm"))

duration <- ggplot(df_phen_event, aes(x = Snowmelt, y = Duration, color = SpeciesID)) +
     #facet_wrap(~ Habitat, nrow=2) +   # a panel for each mountain range
     geom_point(alpha = 0.5) +
     ylab("Duration of activity (No. of days)") +
     xlab("Snowmelt timing (Day of Year)") +
     theme_classic() +
     scale_color_viridis_d(option = "plasma") +
     geom_smooth(se = FALSE, method = "lm", linewidth = 2) +
     theme(legend.position = "none",
           axis.title.y = element_text(vjust = 2, size = 14),
           axis.title.x = element_text(vjust = 0.8, size = 14),
           axis.text.y = element_text(hjust = 0.5, size = 12),
           axis.text.x = element_text(angle = 35, vjust = 0.5, size = 12),
           plot.margin = margin(10,0.5,10,1, "cm"))

ggarrange(peak, duration)

#Visualize community trends
peak_temp <- ggplot(df_phen_event, aes(x = Peak_Temp30, y = Peak, color = SpeciesID)) +
     #facet_wrap(~ Habitat, nrow=2) +   # a panel for each mountain range
     geom_point(alpha = 0.5) +
     ylab("Peak activity (Day of Year)") +
     xlab("Air temperature (ºC)") +
     theme_classic() +
     scale_color_viridis_d(option = "plasma") +
     geom_smooth(se = FALSE, method = "lm", linewidth = 2) +
  #facet_wrap(Habitat ~ .) +
     theme(legend.position = "none",
           axis.title.y = element_text(vjust = 2, size = 14),
           axis.title.x = element_text(vjust = 0.8, size = 14),
           axis.text.y = element_text(hjust = 0.5, size = 12),
           axis.text.x = element_text(angle = 35, vjust = 0.5, size = 12),
           plot.margin = margin(10,1,10,0.5, "cm"))

duration_temp <- ggplot(df_phen_event, aes(x = Onset_Temp30, y = Duration, color = SpeciesID)) +
     #facet_wrap(~ Habitat, nrow=2) +   # a panel for each mountain range
     geom_point(alpha = 0.5) +
     ylab("Duration of activity (No. of days)") +
     xlab("Air temperature (ºC)") +
     theme_classic() +
     scale_color_viridis_d(option = "plasma") +
     geom_smooth(se = FALSE, method = "lm", linewidth = 2) +
     #facet_wrap(Habitat ~ .) +
     theme(legend.position = "none",
           axis.title.y = element_text(vjust = 2, size = 14),
           axis.title.x = element_text(vjust = 0.8, size = 14),
           axis.text.y = element_text(hjust = 0.5, size = 12),
           axis.text.x = element_text(angle = 35, vjust = 0.5, size = 12),
           plot.margin = margin(10,0.5,10,1, "cm"))

ggarrange(peak_temp, duration_temp)
```

**Same models but with no scaling of variables to get information on the rate of phenological response (slope estimates)**

We also want to know how much variance taxon and habitat hold and therefore we compare models with and without random effects:

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.peak <-
  lmer(
    Peak ~ 1 + Snowmelt + Peak_Temp_new + Year + 
      (1 | Plot) +
      (1 + Snowmelt + Peak_Temp_new | speciesplot) +
      (1 + Snowmelt + Peak_Temp_new | SpeciesID) +
      (1 + Snowmelt + Peak_Temp_new | Habitat), REML = TRUE,
    data = df_phen_event
  )

summary(m.peak)

m.peak1 <-
  lmer(
    Peak ~ 1 + Snowmelt + Peak_Temp_new + Year + 
      (1 + Snowmelt + Peak_Temp_new | speciesplot), REML = TRUE,
    data = df_phen_event
  )

summary(m.peak1)

m.peak2 <-
  lmer(
    Peak ~ 1 + Snowmelt + Peak_Temp_new + Year + 
      (1 + Snowmelt + Peak_Temp_new | speciesplot) +
      (1 + Snowmelt + Peak_Temp_new | SpeciesID), REML = TRUE,
    data = df_phen_event
  )

summary(m.peak2)

m.peak3 <-
  lmer(
    Peak ~ 1 + Snowmelt + Peak_Temp_new + Year + 
      (1 | Plot) +
      (1 + Snowmelt + Peak_Temp_new | speciesplot) +
      (1 + Snowmelt + Peak_Temp_new | Habitat), REML = TRUE,
    data = df_phen_event
  )

summary(m.peak3)

AIC(m.peak, m.peak1, m.peak2, m.peak3)

peak_coefs <- summary(m.peak)$coefficients[2,1:ncol(summary(m.peak)$coefficients)]
peak_coefs <- data.frame(summary(m.peak)$coefficients[c(2:3),c(1:3)])
peak_coefs$variable <- c("Snowmelt", "Temperature")
peak_coefs$pheno <- "Peak"

```

**Table S5.3** To test the variance partitioned between taxa and plot in peak phenological response to snowmelt timing and temperature, we compared the full model (including sampling variance, taxa and plot) with simple models including one random effect.

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.peak, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Peak ~ Snowmelt + Temp + Year + (1 + Temp + snow | Taxa) + (1 + Temp + snow | Plot)", terms = c("(Intercept)", "Snowmelt", "Peak_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.peak1, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Peak ~ Snowmelt + Temp + Year", terms = c("(Intercept)", "Snowmelt", "Peak_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.peak2, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Peak ~ Snowmelt + Temp + Year + (1 + Temp + snow | Taxa)", terms = c("(Intercept)", "Snowmelt", "Peak_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.peak3, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Peak ~ Snowmelt + Temp + Year + (1 + Temp + snow | Habitat)", terms = c("(Intercept)", "Snowmelt", "Peak_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

Coefficients from full model

```{r}
coef_peak <- coef(m.peak)$SpeciesID

library(knitr)

coefficients_df <- as.data.frame(coef_peak)

coefficients_table <- kable(coefficients_df, caption = "Coefficients Table")

print(coefficients_table)

```

```{r}
coef(m.peak)$Habitat
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.duration <-
  lmer(
    Duration ~ 1 + Snowmelt + Onset_Temp_new + Year +
      (1 | Plot) +
      (1 + Snowmelt + Onset_Temp_new | speciesplot) +
      (1 + Snowmelt + Onset_Temp_new | SpeciesID) +
      (1 + Snowmelt + Onset_Temp_new | Habitat),
    data = df_phen_event
  )

summary(m.duration)

m.duration1 <-
  lmer(
    Duration ~ 1 + Snowmelt + Onset_Temp_new + Year + 
      (1 + Snowmelt + Onset_Temp_new | speciesplot),
    data = df_phen_event
  )

summary(m.duration1)

m.duration2 <-
  lmer(
    Duration ~ 1 + Snowmelt + Onset_Temp_new + Year + 
      (1 + Snowmelt + Onset_Temp_new | speciesplot) +
      (1 + Snowmelt + Onset_Temp_new | SpeciesID),
    data = df_phen_event
  )

summary(m.duration2)

m.duration3 <-
  lmer(
    Duration ~ 1 + Snowmelt + Onset_Temp_new + Year +
      (1 | Plot) +
      (1 + Snowmelt + Onset_Temp_new | speciesplot) +
      (1 + Snowmelt + Onset_Temp_new | Habitat),
    data = df_phen_event
  )

summary(m.duration3)

duration_coefs <- summary(m.duration)$coefficients[2,1:ncol(summary(m.duration)$coefficients)]
duration_coefs <- data.frame(summary(m.duration)$coefficients[c(2:3),c(1:3)])
duration_coefs$variable <- c("Snowmelt", "Temperature")
duration_coefs$pheno <- "Duration"
```

**Table S5.4** To test the variance partitioned between taxa and plot in duration of phenology to snowmelt timing and temperature, we compared the full model (including sampling variance, taxa and plot) with simple models including one random effect.

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.duration, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Duration ~ Snowmelt + Temp + Year + (1 + Temp + snow | Taxa) + (1 + Temp + snow | Plot)", terms = c("(Intercept)", "Snowmelt", "Onset_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.duration1, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Duration ~ Snowmelt + Temp + Year", terms = c("(Intercept)", "Snowmelt", "Onset_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.duration2, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Duration ~ Snowmelt + Temp + Year + (1 + Temp + snow | Taxa)", terms = c("(Intercept)", "Snowmelt", "Onset_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
#| message: false
#| warning: false
#| echo: false

tab_model(m.duration3, show.intercept = TRUE, show.est = TRUE, show.se = TRUE, show.p = TRUE, show.stat = TRUE, show.ci = FALSE,
         title = "Duration ~ Snowmelt + Temp + Year + (1 + Temp + snow | Plot)", terms = c("(Intercept)", "Snowmelt", "Onset_Temp_new", "Year"), pred.labels = c("Intercept", "Snowmelt", "Temp", "Year"), dv.labels = "", string.est = "Coeff", string.se = "SE", string.stat = "t", string.p = "P-Value")
```

```{r}
coef(m.duration)$SpeciesID
```

```{r}
coef(m.duration)$Habitat
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.onset <-
  lmer(
    Onset ~ 1 + Snowmelt + Onset_Temp_new + Year + 
      (1 | Plot) +
      (1 + Snowmelt + Onset_Temp_new | speciesplot) +
      (1 + Snowmelt + Onset_Temp_new | SpeciesID) +
      (1 + Snowmelt + Onset_Temp_new | Habitat),
    data = df_phen_event
  )

summary(m.onset)

onset_coefs <- summary(m.onset)$coefficients[2,1:ncol(summary(m.onset)$coefficients)]
onset_coefs <- data.frame(summary(m.onset)$coefficients[c(2:3),c(1:3)])
onset_coefs$variable <- c("Snowmelt", "Temperature")
onset_coefs$pheno <- "Onset"
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

m.end <-
  lmer(
    End ~ 1 + Snowmelt + End_Temp_new + Year +
      (1 | Plot) +
      (1 + Snowmelt + End_Temp_new | speciesplot) +
      (1 + Snowmelt + End_Temp_new | SpeciesID) +
      (1 + Snowmelt + End_Temp_new | Habitat),
    data = df_phen_event
  )

summary(m.end)

end_coefs <- summary(m.end)$coefficients[2,1:ncol(summary(m.end)$coefficients)]
end_coefs <- data.frame(summary(m.end)$coefficients[c(2:3),c(1:3)])
end_coefs$variable <- c("Snowmelt", "Temperature")
end_coefs$pheno <- "End"
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

df_combine <- bind_rows(onset_coefs,peak_coefs,end_coefs,duration_coefs)

df_combine <-
  readxl::read_xlsx("Data/phenology_data/df_combine.xlsx")

```

```{r}
#| message: false
#| warning: false
#| echo: false

df_combine$pheno <- factor(df_combine$pheno,                 # Relevel group factor
                         levels = c("Onset", "Peak", "End", "Duration"))

ggplot(df_combine, aes(pheno, Estimate, fill = pheno)) +
  xlab("") +
  ylab("Phenological response") +
  coord_cartesian(ylim=c(-4, 4)) +
  geom_errorbar(aes(ymin = Estimate - Std..Error, ymax = Estimate + Std..Error, color = pheno), alpha = 0.8, width = 0.1, linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 3, shape = 21, color = "black") +
  facet_grid(. ~ variable) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") +
  theme_bw() +
  theme(axis.text.y = element_text(hjust = 0.5, size = 12),
      axis.text.x = element_text(angle = 35, vjust = 0.5, size = 12),
      axis.title.y = element_text(vjust = 2, size = 14))
    #plot.margin = margin(10,4,10,4,"cm"))
```

**Figure S5.2** Arthropod community phenological responses to snowmelt timing and temperature across all arthropod phenological events (onset, peak, end and duration of activity).
